---
title: "LC50 Nickel Experiment 1"
author: "Celeste Valdivia"
date: "2024-04-10"
output: html_document
---

```{r, eval =TRUE}
library(knitr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(hrbrthemes)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE)
```

# Retrieving Data from Google Sheets


Make sure you have made your Google sheet publicly available to anyone that has the link. If you make any updates to the sheet just re-curl the data, meaning just re-run the code below.

```{r, engine='bash', eval=FALSE}
cd ..

curl -L https://docs.google.com/spreadsheets/d/1JtJO3EX06BYK4pYwZZ7b7ZCoTyZcAeuTRka_9kA-yPk/export?exportFormat=csv | tee data/LC50.csv
```

Read in the data to your local R environment.

```{r, eval=TRUE}
setwd('..')
LC50 <- read.csv(file = "data/LC50.csv")  
```

# Cleaning up Data

```{r, eval=TRUE}
LC50$Date <- mdy(LC50$Date) #convert the Date column from characters to true Date

# Create a new column 'simple_Stage' based on conditions
LC50 <- LC50 %>%
  mutate(simple_Stage = case_when(
    Stage %in% c("A1", "A2") ~ "A",
    Stage %in% c("B1", "B2") ~ "B",
    Stage %in% c("C1", "C2") ~ "C",
    Stage == "D" ~ "D",
    TRUE ~ NA_character_  # This will handle any other cases or return NA if none match
  ))

# Create new simplified health scoring
LC50 <- LC50 %>%
  mutate(simple_health = case_when(
    Health == "0" ~ "0",
    Health %in% c("1", "2") ~ "1",
    Health %in% c("3", "4") ~ "2",
    Health %in% c("5", "6") ~ "3",
    Health %in% c("7", "8") ~ "4",
    Health %in% c("9", "10") ~ "5",
    TRUE ~ NA_character_)) %>%
  transform(simple_health = as.numeric(simple_health))

# add a total column
summary(LC50)
```

Will need to exclude two individuals from study. DM032024_C03 (0.1 mg/L condition) and DM042024_C02 (1 mg/L condition). Results in 8 replicates for all treatments except 7 only for the 1 mg/L condition.

DM032024_C03: Cut at 24 hour mark.

DM042024_C02: Decline in health as it was not healthy enough to begin with.

```{r}
LC50 <- LC50[LC50$Tunicate.ID != c("DM032024_C03", "DM042024_C02"),]
```


# Summary Stats

```{r}
avg_zooid <- mean(LC50$Total.Zooids, na.rm = TRUE)
sd_zooid <- sd(LC50$Total.Zooids, na.rm = TRUE)
print(c(avg_zooid, sd_zooid))
```


# Calculate Percent Mortality per Treatment

```{r}
calc_mort_perecnt <- function(data) {
  
  total_individuals <- aggregate(Mortality ~ hpe + Ni.Conc., data = data, FUN = length)
  
  total_mortality <- aggregate(Mortality ~ hpe + Ni.Conc., data = data, FUN = sum)
  
  merged_data <- merge(total_individuals, total_mortality, by = c( "hpe",  "Ni.Conc."), suffixes = c(".total", ".count"))
  
  merged_data$percentage_mortality <- (merged_data$Mortality.count / merged_data$Mortality.total) *100
  
  return(merged_data)
}

mortality_percentage <- calc_mort_perecnt(LC50)

mortality_percentage$hpe <- as.character(mortality_percentage$hpe) # Did this so I can have discrete colors for plot for the time points instead of a gradient

print(mortality_percentage)
summary(mortality_percentage)

hist(mortality_percentage$percentage_mortality)

```
Data follows a bimodal distribution.


# Calculate Survivorship
```{r}
calc_survivor_percent <- function(data) {
  
  total_individuals <- aggregate(Alive ~ hpe + Ni.Conc., data = data, FUN = length)
  
  total_survivor <- aggregate(Alive ~ hpe + Ni.Conc., data = data, FUN = sum)
  
  merged_data <- merge(total_individuals, total_survivor, by = c( "hpe",  "Ni.Conc."), suffixes = c(".total", ".count"))
  
  merged_data$percentage_survivor <- (merged_data$Alive.count / merged_data$Alive.total) *100
  
  return(merged_data)
}

survival_percent <- calc_survivor_percent(LC50)

survival_percent$hpe <- as.character(survival_percent$hpe) # Did this so I can have discrete colors for plot for the time points instead of a gradient

print(survival_percent)
summary(survival_percent)

survival_percent$Ni.Conc. <- as.numeric(survival_percent$Ni.Conc.) # Did this so geom_smooth actually works
```




# Exploratory percent mort graph

```{r}
mort_exp <- subset(mortality_percentage, hpe != 0)


# Plot the data with smoothed curve
ggplot(mort_exp, aes(x = Ni.Conc., y = percentage_mortality)) + 
  geom_point(size = 2) +
  facet_wrap(~ hpe) +
  geom_smooth(method = "loess", se = TRUE, span = 0.5, color = "black") + # the geom smooth method is currently incorrect, need to find out how to display what was found to be the LC50 using LC_logit from the ecotox package
  geom_hline(yintercept = 50, linetype = "dotted", color = "red") +  # Add red dotted line at y = 50
  labs(x = "Nickel Concentration mg/L",
       y = "Mortality (%)")

# Plot the data with smoothed curve
ggplot(mort_exp, aes(x = Ni.Conc., y = percentage_mortality)) + 
  geom_point(size = 2) +
  facet_wrap(~ hpe) +
geom_line() +
  geom_hline(yintercept = 50, linetype = "dotted", color = "red") +  # Add red dotted line at y = 50
  labs(x = "Nickel Concentration mg/L",
       y = "Mortality (%)")

```

Exploratory survival graph
```{r}
ggplot(survival_percent, aes(x= Ni.Conc., y=percentage_survivor, color = hpe)) + 
  facet_wrap(~hpe) +
  geom_point() +
  geom_line() +
      labs(x = "Nickel mg/L",
       y = "Survival (%)")  
```




# Dose Response Models Readings

We do not have a traditional sigmoidal dose response curve as an option. Botryllus remain with heartbeat up to 100 mg/L and cease blood movement at 1000 mg/L after 48 hours.

https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0146021

https://www.linkedin.com/pulse/simple-lc50-calculation-pesticide-toxicity-bioassay-using-adhiwibawa

Eventually will need to explore packages for assessing LC50 or EC50.

Also need to decide that for the ecotox modeling if it makes sense to use a probit, logit, or linear regression model to assess LC50 for whatever time point of interest.

Probit Model:
The probit model assumes that the dose-response relationship follows a cumulative *normal distribution*.
It is often used when the response variable is binary (alive/dead) or when the response follows a sigmoidal shape.
Suitable when responses are discrete and there's a natural threshold for response.

Logit Model:
Similar to the probit model, the logit model is used for binary response data.
It assumes that the log odds of the response is linearly related to the dose.
Like probit, it's suitable for binary responses but may be preferred for computational reasons or if assumptions align better with the data.

Linear Model:
The linear model assumes a linear relationship between the dose and the response.
It may be appropriate when responses are continuous and show a linear trend across doses.
However, it may not capture non-linear relationships well, especially if the response plateaus or changes direction.

Given my data follows a bimodal distribution it would most likely be most appropriate to follow a logit model.

```{r, dummy ecotox data}
#install.packages("ecotox")
library(ecotox)

head(lamprey_tox)

hist(lamprey_tox$response)

m <- LC_logit((response / total) ~ log10(dose), p = c(50, 99),
weights = total,
data = lamprey_tox[lamprey_tox$nominal_dose != 0, ],
subset = c(month == "May"))

print(m)

lc_may <- subset(lamprey_tox, month %in% c("May"))
p1 <- ggplot(data = lc_may[lc_may$nominal_dose != 0, ],
  aes(x = dose, y = (response / total))) +
  geom_point() +
  geom_smooth(method = "glm",
   method.args = list(family = binomial(link = "logit")),
 aes(weight = total), colour = "#FF0000", se = TRUE)

 p1

```

```{r}
head(LC50)

LC50_24h <- LC_logit((Mortality / Total) ~ log10(Ni.Conc.), p = c(50),
weights = Total,
data = LC50[LC50$Ni.Conc. != 0, ],
subset = c(hpe == "24"))

print(LC50_24h)
```

```{r}
LC50_48h <- LC_logit((Mortality / Total) ~ log10(Ni.Conc.), p = c(50),
weights = Total,
data = LC50[LC50$Ni.Conc. != 0, ],
subset = c(hpe == "48"))

print(LC50_48h)
```

```{r}
LC50_72h <- LC_logit((Mortality / Total) ~ log10(Ni.Conc.), p = c(50),
weights = Total,
data = LC50[LC50$Ni.Conc. != 0, ],
subset = c(hpe == "72"))

print(LC50_72h)
```
```{r}
LC50_96h <- LC_logit((Mortality / Total) ~ log10(Ni.Conc.), p = c(50),
weights = Total,
data = LC50[LC50$Ni.Conc. != 0, ],
subset = c(hpe == "96"))
print(LC50_96h)
```


combine tibbles
```{r}
library(dplyr)

# Assuming LC50_24, LC50_48, ..., LC50_96 are tibbles/data frames containing LC50 results
# Each tibble should have at least two columns: 'dose' and 'LC50'

# Combine the LC50 results into one data frame
combined_lc50 <- bind_rows(
  LC50_24h %>% mutate(time_point = 24),
  LC50_48h %>% mutate(time_point = 48),
  LC50_72h %>% mutate(time_point = 72),
  LC50_96h %>% mutate(time_point = 96)
)

# Now you have a single data frame 'combined_lc50' containing LC50 results for different time points
# It should have columns: 'dose', 'LC50', and 'time_point'
```

```{r}
# You can now plot the LC50 values over time
#install.packages("ggrepel")
library(ggrepel)

#setwd('..')
#png(filename = "output/LC50_overtime.png", width = 500, height = 300)

# You can now plot the LC50 values over time
ggplot(combined_lc50, aes(x = time_point, y = dose)) +
  geom_point() +
  geom_smooth(method = "loess", color = "black") +  # Set line color to black
  geom_label_repel(aes(label = paste(time_point,"hour", "LC50", "=", round(dose), "mg/L")), 
                   size = 3, 
                   nudge_x = 2, 
                   nudge_y = 2,
                   arrow = arrow(length = unit(2, "npc"))) +  # Add labels with arrows
  labs(x = "Time (hours)", y = " Nickel Concentration (mg/L)") +
  ylim(0, 1000)  # Set y-axis limits

#dev.off()

```



# Exploratory Blastogenic Arrest

https://www.sciencedirect.com/science/article/pii/S0012160618303828

above they refer to arrest and they interchangibly refer to torpor in 

We may define effect concentration 50% by blastogenic arrest or blastogenic development inhibition in the animal as a chronic endpoint of nickel toxicity. 

a. Need to account for rate of blastogenesis again. One step per day is normal and should be found in the control animal.

b. In the 100 treatment for more than 3 days we see no change in blastogenic rate. The line will be flat over time.

Exploratory health
```{r}
ggplot(LC50, aes(x = hpe, y = simple_health, color = Ni.Conc.)) + 
  geom_point() +
  facet_wrap(~ Ni.Conc., labeller = as_labeller(c("0" = "0 mg/L", "0.1" = "0.1 mg/L", "1" = "1 mg/L", "10" = "10 mg/L", "100" = "100 mg/L", "1000" = "1000 mg/L"))) +
  scale_x_continuous(breaks = c(0, 24, 48, 72, 96)) +
  ylim(0, 5) +
  geom_smooth(aes(fill = Ni.Conc.), method = "lm", se = TRUE) +  # Add linear regression line with shaded SE
  labs(x = "hpe", 
       y = "Health"
       )

```


ddfff
```{r}
y_breaks <- 1:7
y_labels <- c("A1", "A2", "B1", "B2", "C1", "C2", "D")

stage_mapping <- c(A1 = 1, A2 = 2, B1 = 3, B2 = 4, C1 = 5, C2 = 6, D = 7)

LC50$Stage_Num <- stage_mapping[LC50$Stage]
```



```{r}
ggplot(LC50, aes(x = hpe, y = Stage_Num)) + 
  geom_point(alpha = 0.6, size = 2) +  # Adjust transparency and size of points
  geom_line(aes(group = Tunicate.ID)) +  # Connect points with lines for each Tunicate.ID
  scale_y_continuous(breaks = y_breaks, labels = y_labels) +  # Custom breaks and labels for y-axis
  scale_x_continuous(breaks = c(0, 24, 48, 72, 96)) +  # Custom breaks for x-axis
  labs(x = "hpe", y = "stage") +  # Axis labels
  facet_wrap(~ Ni.Conc., labeller = as_labeller(c("0" = "0 mg/L", "0.1" = "0.1 mg/L", "1" = "1 mg/L", "10" = "10 mg/L", "100" = "100 mg/L", "1000" = "1000 mg/L")))
```


# Test Data

```{r}
df <- LC50[LC50$Tunicate.ID == "DM012024_C01", ]

stage_mapping <- c(A1 = 1, A2 = 2, B1 = 3, B2 = 4, C1 = 5, C2 = 6, D = 7)

df$Stage_Num <- stage_mapping[df$Stage]

summary(df)
```

```{r}
y_breaks <- 1:7
y_labels <- c("A1", "A2", "B1", "B2", "C1", "C2", "D")

ggplot(df, aes(x = hpe, y = Stage_Num, color = Ni.Conc.)) + 
  geom_point(alpha = 0.6, size = 3) +  # Adjust transparency and size of points
  geom_smooth(method = "loess", se = FALSE, span = 0.9)  +
  scale_y_continuous(breaks = y_breaks, labels = y_labels) +  # Custom breaks and labels for y-axis
    scale_x_continuous(breaks = c(0, 24, 48, 72, 96)) +
  labs(x = "hpe", y = "stage")

```