---
title: "02.2-LC50-ammendment"
output: html_document
date: "2025-08-12"
---

# Setup

```{r, eval =TRUE}
library(knitr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(hrbrthemes)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(ecotox)
library(wesanderson)
library(viridis)
library(ggrepel)
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE)
```

# Retrieving Data from Google Sheets

Make sure you have made your Google sheet publicly available to anyone that has the link. If you make any updates to the sheet just re-curl the data, meaning just re-run the code below.

## Mortality Data

```{r, engine='bash', eval=FALSE}
cd ..

curl -L https://docs.google.com/spreadsheets/d/1JtJO3EX06BYK4pYwZZ7b7ZCoTyZcAeuTRka_9kA-yPk/export?exportFormat=csv | tee data/LC50.csv
```

## Nickel ICPMS Data for LC50 estimation

```{r, engine='bash', eval=FALSE}
cd ..

curl -L https://docs.google.com/spreadsheets/d/1sCogpUsK6rZ7--VgcameLrresiCdPaMDBOaPZvJWWHg/export?exportFormat=csv | tee data/icpms.ni.csv
```

Read in the data to your local R environment.

```{r, eval=TRUE}
setwd('..')
LC50 <- read.csv(file = "data/LC50.csv")  
icpms <- read.csv(file = "data/icpms.ni.csv")
```

# Cleaning up Data

## Add simplified scoring columns and convert dates

Convert Dates

```{r, eval=TRUE}
LC50$date <- mdy(LC50$date) #convert the Date column from characters to true Date
```

Simple Stage column added:

```{r}
LC50 <- LC50 %>%
  mutate(simple_stage = case_when(
    stage %in% c("A1", "A2") ~ "A",
    stage %in% c("B1", "B2") ~ "B",
    stage %in% c("C1", "C2") ~ "C",
    stage == "D" ~ "D",
    TRUE ~ NA_character_  # This will handle any other cases or return NA if none match
  ))
```

Simple health scoring column added

```{r}
# Create new simplified health scoring
LC50 <- LC50 %>%
  mutate(simple_health = case_when(
    Health == "0" ~ "0",
    Health %in% c("1", "2") ~ "1",
    Health %in% c("3", "4") ~ "2",
    Health %in% c("5", "6") ~ "3",
    Health %in% c("7", "8") ~ "4",
    Health %in% c("9", "10") ~ "5",
    TRUE ~ NA_character_)) %>%
  transform(simple_health = as.numeric(simple_health))

summary(LC50)
```

## Excluding tunicate outliers

Will need to exclude two individuals from study. DM032024_C03 (0.1 mg/L condition) and DM042024_C02 (1 mg/L condition). Results in 8 replicates for all treatments except 7 only for the 1 mg/L condition.

DM032024_C03: Cut at 24 hour mark.

DM042024_C02: Decline in health as it was not healthy enough to begin with.

DM022024_C01: Does not have a match for the SOD assay at the 96 hr mark so will be excluded from that. Will need to be consistent.

```{r}
t_LC50 <- LC50[!is.na(LC50$true_rep), ]
```

##

```{r}
icpms <- icpms[icpms$experiment != "DM",] # exclude the De Moines environmental water samples confirming absence of nickel in environment from which tunicates were pulled 
```



# Survivorship

For now will not be using Survivorship in final deliverable but may be useful in future toxicology studies

## Survivorship Function

```{r}
calc_survivor_percent <- function(data) {
  
  total_individuals <- aggregate(Alive ~ hpe + NiCl2.Nom.Conc., data = data, FUN = length)
  
  total_survivor <- aggregate(Alive ~ hpe + NiCl2.Nom.Conc., data = data, FUN = sum)
  
  merged_data <- merge(total_individuals, total_survivor, by = c( "hpe",  "NiCl2.Nom.Conc."), suffixes = c(".total", ".count"))
  
  merged_data$percentage_survivor <- (merged_data$Alive.count / merged_data$Alive.total) *100
  
  return(merged_data)
}

survival_percent <- calc_survivor_percent(t_LC50)

survival_percent$hpe <- as.character(survival_percent$hpe) # Did this so I can have discrete colors for plot for the time points instead of a gradient

print(survival_percent)
summary(survival_percent)

survival_percent$NiCl2.Nom.Conc. <- as.numeric(survival_percent$NiCl2.Nom.Conc.) # Did this so geom_smooth actually works
```

# Survivor Graphs

### subset data

```{r}
surv_24_96 <-subset(survival_percent, hpe %in% c(24, 96))
```

for nominal nickel(II) chloride calculating 50% crude mortality lc50
```{r}
# Calculate the slope and intersection points for each hpe group
fit_data <- surv_24_96 %>%
  filter(NiCl2.Nom.Conc. %in% c(100, 1000)) %>%
  group_by(hpe) %>%
  summarize(
    slope = (percentage_survivor[NiCl2.Nom.Conc. == 1000] -     percentage_survivor[NiCl2.Nom.Conc. == 100]) / 
            (1000 - 100),
    intercept = percentage_survivor[NiCl2.Nom.Conc. == 100] - slope * 100,
    x_intercept50 = (50 - intercept) / slope,
    x_intercept90 = (90 - intercept) / slope,
    equation = paste("y = ", round(slope, 2), "x + ", round(intercept, 2))

  )

# Merge the intersection data with the original data
surv_24_96 <- surv_24_96 %>%
  left_join(fit_data, by = "hpe")
```
### survival 24 and 96 hpe


nominal nickel (II) chloride
```{r}
setwd('..')
png(filename = "output/survival_24_96_2025_nofacet.png", width = 900, height = 700)

ggplot(surv_24_96, aes(x = NiCl2.Nom.Conc., y = percentage_survivor, color = factor(hpe))) + 
  
  # Points with different shapes for 24h vs 96h
  geom_point(aes(shape = factor(hpe)), size = 5) +

  # Lines connecting points
  geom_line(aes(group = hpe), size = 2) +

  # Manual color scale
  scale_color_manual(values = c("lightblue", "darkblue"), name = "hpe") +

  # Manual shape scale: solid for 24, open for 96
  scale_shape_manual(values = c("24" = 16, "96" = 18), name = "hpe") +

  labs(x = "Nominal Nickel (II) Chloride (mg/L)",
       y = "Survival (%)") +

  scale_y_continuous(breaks = seq(0, 100, by = 10), expand = c(0.02, 0.02)) +
  scale_x_continuous(expand = c(0.02, 0.02)) +

  theme_test() +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    axis.text.y = element_text(size = 20, hjust = 1.2, margin = margin(r = 20)),
    axis.text.x = element_text(size = 20, vjust = 3, hjust = 0.5, margin = margin(t = 20)),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    strip.text = element_blank(),  
    strip.background = element_blank(),
    panel.spacing = unit(5, "lines")
  ) +

  # Labels for 50% survival
 # geom_label_repel(
  #  data = fit_data,
 #   aes(x = x_intercept50, y = 50, label = paste(round(x_intercept50), "mg/L")),
 #   box.padding = unit(2, "lines"),
 #   label.padding = unit(0.6, "lines"),
  #  point.padding = unit(0.5, "lines"),
 #   segment.size = 1,
#    size = 8,
#    color = "black",
#    nudge_x = 50,
 #   nudge_y = 5,
 
  #   show.legend = FALSE
#  ) +

  # 50% survival lines
  geom_segment(data = fit_data, 
               aes(x = min(surv_24_96$NiCl2.Nom.Conc.), 
                   xend = x_intercept50, 
                   y = 50, yend = 50), 
               linetype = "dashed", color = "red3", linewidth = 1) +  
  geom_segment(aes(x = x_intercept50, xend = x_intercept50, 
                   y = 0, yend = 50), 
               linetype = "dashed", color = "red3", linewidth = 1)

dev.off()
```

# Nominal Ni Concentrations
## Create new data frame for calculated values

Calculate the fraction of ni in anhydrous nickel (II) chloride
```{r}
# atomic weights
ni_aw <- 58.6934
cl_aw <- 35.453

# molar mass of NiCl2 (g/mol)
nicl2_mw <- ni_aw + (2 * cl_aw)

# fraction of Ni in NiCl2
ni_fraction <- ni_aw / nicl2_mw
ni_fraction

```
Calculates the amount of nominal nickel in the nominal concentrations of nickel(II) chloride anhydrous
```{r}
ni_conversion <- surv_24_96 %>%
  dplyr::distinct(NiCl2.Nom.Conc.) %>%
  dplyr::mutate(
    Ni_mg_L = NiCl2.Nom.Conc. * ni_fraction
  )
```
```{r}
surv_24_96 <- surv_24_96 %>%
  dplyr::left_join(ni_conversion, by = "NiCl2.Nom.Conc.")
```

```{r}
# Calculate the slope and intersection points for each hpe group
x1 <- 45.28832695 # calculated amount of ni in the 100 treatment
x2 <- 452.88326952 # calculated amount of ni in the 1000 treatment

fit_data_ni <- surv_24_96 %>%
  mutate(Ni_mg_L_r = round(Ni_mg_L, 6)) %>%
  filter(Ni_mg_L_r %in% round(c(x1, x2), 6)) %>%
  group_by(hpe) %>%
  summarise(
    y1 = percentage_survivor[Ni_mg_L_r == round(x1, 6)],
    y2 = percentage_survivor[Ni_mg_L_r == round(x2, 6)],
    slope_ni = (y2 - y1) / (x2 - x1),
    intercept_ni = y1 - slope_ni * x1,
    x_intercept50_ni = (50 - intercept_ni) / slope_ni,
    x_intercept90_ni = (90 - intercept_ni) / slope_ni,
    equation_ni = paste0("y = ", round(slope_ni, 3), "x + ", round(intercept_ni, 2)),
    .groups = "drop"
  )


# Merge the intersection data with the original data
surv_24_96 <- surv_24_96 %>%
  left_join(fit_data_ni, by = "hpe")
```


## nom ni plot
```{r}
setwd('..')
png(filename = "output/nom_ni_survival_24_96_2025_nofacet.png", width = 900, height = 700)

ggplot(surv_24_96, aes(x = Ni_mg_L, y = percentage_survivor, color = factor(hpe))) + 
  
  # Points with different shapes for 24h vs 96h
  geom_point(aes(shape = factor(hpe)), size = 5) +

  # Lines connecting points
  geom_line(aes(group = hpe), size = 2) +

  # Manual color scale
  scale_color_manual(values = c("lightblue", "darkblue"), name = "hpe") +

  # Manual shape scale: solid for 24, open for 96
  scale_shape_manual(values = c("24" = 16, "96" = 18), name = "hpe") +

  labs(x = "Nominal Nickel (mg/L)",
       y = "Survival (%)") +

  scale_y_continuous(breaks = seq(0, 100, by = 10), expand = c(0.02, 0.02)) +
  scale_x_continuous(expand = c(0.02, 0.02)) +

  theme_test() +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    axis.text.y = element_text(size = 20, hjust = 1.2, margin = margin(r = 20)),
    axis.text.x = element_text(size = 20, vjust = 3, hjust = 0.5, margin = margin(t = 20)),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    strip.text = element_blank(),  
    strip.background = element_blank(),
    panel.spacing = unit(5, "lines")
  ) +

  # Labels for 50% survival
 # geom_label_repel(
  #  data = fit_data,
 #   aes(x = x_intercept50, y = 50, label = paste(round(x_intercept50), "mg/L")),
 #   box.padding = unit(2, "lines"),
 #   label.padding = unit(0.6, "lines"),
  #  point.padding = unit(0.5, "lines"),
 #   segment.size = 1,
#    size = 8,
#    color = "black",
#    nudge_x = 50,
 #   nudge_y = 5,
 
  #   show.legend = FALSE
#  ) +

  # 50% survival lines
  geom_segment(data = fit_data_ni, 
               aes(x = min(surv_24_96$Ni_mg_L), 
                   xend = x_intercept50_ni, 
                   y = 50, yend = 50), 
               linetype = "dashed", color = "red3", linewidth = 1) +  
  geom_segment(aes(x = x_intercept50_ni, 
                   xend = x_intercept50_ni, 
                   y = 0, yend = 50), 
               linetype = "dashed", color = "red3", linewidth = 1)

dev.off()
```

# ICPMS measured Ni
## create new data frame
```{r}
icpms_summary <- icpms %>%
  group_by(NiCl2.Nom.Conc.) %>%
  summarise(
    Ni_mean_mg_L = mean(Ni.mg.L, na.rm = TRUE),
    Ni_sd_mg_L   = sd(Ni.mg.L, na.rm = TRUE),
    Ni_min_mg_L  = min(Ni.mg.L, na.rm = TRUE),
    Ni_max_mg_L  = max(Ni.mg.L, na.rm = TRUE),
    n_icp        = sum(!is.na(Ni.mg.L)),
    .groups = "drop"
  )



# 2) Join onto survival dataframe
surv_24_96_icpms <- surv_24_96 %>%
  left_join(icpms_summary, by = "NiCl2.Nom.Conc.")
 
surv_24_96_icpms %>%
  distinct(NiCl2.Nom.Conc., Ni_mean_mg_L, Ni_min_mg_L, Ni_max_mg_L, n_icp) %>%
  arrange(NiCl2.Nom.Conc.)

```

```{r}
fit_data_measNi <- surv_24_96_icpms %>%
  filter(NiCl2.Nom.Conc. %in% c(100, 1000)) %>%
  group_by(hpe) %>%
  summarise(
    x1 = Ni_mean_mg_L[NiCl2.Nom.Conc. == 100],
    x2 = Ni_mean_mg_L[NiCl2.Nom.Conc. == 1000],
    y1 = percentage_survivor[NiCl2.Nom.Conc. == 100],
    y2 = percentage_survivor[NiCl2.Nom.Conc. == 1000],

    slope_measNi = (y2 - y1) / (x2 - x1),
    intercept_measNi = y1 - slope_measNi * x1,

    x_intercept50_measNi = (50 - intercept_measNi) / slope_measNi,
    x_intercept90_measNi = (90 - intercept_measNi) / slope_measNi,

    equation_measNi = paste0(
      "y = ", round(slope_measNi, 3), "x + ", round(intercept_measNi, 2)
    ),
    .groups = "drop"
  )

```

```{r}
setwd('..')
png(filename = "output/icpms_ni_survival_24_96_2025_nofacet.png", width = 900, height = 700)

ggplot(surv_24_96_icpms, aes(x = Ni_mean_mg_L, y = percentage_survivor, color = factor(hpe))) + 
  
  # (Optional) show ICP-MS range for each nominal group as horizontal bars
  # Comment out if you don't want it
  #geom_errorbarh(
  #  aes(xmin = Ni_min_mg_L, xmax = Ni_max_mg_L),
#    height = 0, linewidth = 1, alpha = 0.7
 # ) +

  # Points with different shapes for 24h vs 96h
  geom_point(aes(shape = factor(hpe)), size = 5) +

  # Lines connecting points
  geom_line(aes(group = hpe), size = 2) +

  # Manual color scale
  scale_color_manual(values = c("lightblue", "darkblue"), name = "hpe") +

  # Manual shape scale
  scale_shape_manual(values = c("24" = 16, "96" = 18), name = "hpe") +

  labs(
    x = "Nickel-60 (mg/L), Measured with ICP-MS",
    y = "Colony Survival (%)"
  ) +

  scale_y_continuous(breaks = seq(0, 100, by = 10), expand = c(0.02, 0.02)) +
  scale_x_continuous(breaks = seq(0, 300, by = 25), expand = c(0.02, 0.02)) +

  theme_test() +
  theme(
    axis.ticks = element_blank(),
    axis.line = element_line(colour = "grey50"),
    panel.grid = element_line(color = "white"),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.background = element_rect(fill = "white", color = "white"),
    plot.background = element_rect(fill = "white", color = "white"),
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    axis.text.y = element_text(size = 20, hjust = 1.2, margin = margin(r = 20)),
    axis.text.x = element_text(size = 20, vjust = 3, hjust = 0.5, margin = margin(t = 20)),
    legend.text = element_text(size = 20),
    legend.title = element_text(size = 20),
    strip.text = element_blank(),  
    strip.background = element_blank(),
    panel.spacing = unit(5, "lines")
  ) +

  # 50% survival horizontal line: from min measured Ni to the 50% x-intercept (per hpe)
  geom_segment(
    data = fit_data_measNi,
    aes(
      x = min(surv_24_96_icpms$Ni_mean_mg_L, na.rm = TRUE),
      xend = x_intercept50_measNi,
      y = 50, yend = 50
    ),
    linetype = "dashed", color = "red3", linewidth = 1,
    inherit.aes = FALSE
  ) +

  # 50% survival vertical line at the x-intercept (per hpe)
  geom_segment(
    data = fit_data_measNi,
    aes(
      x = x_intercept50_measNi, xend = x_intercept50_measNi,
      y = 0, yend = 50
    ),
    linetype = "dashed", color = "red3", linewidth = 1,
    inherit.aes = FALSE
  )
dev.off()
```


