---
title: "03-RNAseqSamples"
author: "Celeste Valdivia"
date: "2024-04-15"
output: html_document
---

```{r, eval =TRUE}
library(knitr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(hrbrthemes)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(stringr)
knitr::opts_chunk$set(echo = TRUE,
                      eval = FALSE)
```

# Objective

Determine which 12 of the 39 RNA samples available from the Autumn 2023 nickel trials to send off for RNAseq. We are targeting all late stage (C2-D) and high (45 mg/L nickel) vs control.

# Revtrieve Data
We will need both the morphometric data and the nanodrop readings. 

```{r, engine='bash', eval=FALSE}
cd ..

curl -L https://docs.google.com/spreadsheets/d/10uM3N3PD9xIP4yUnadfhkcXa8TPMRMD-adOIKbmkYzY/export?exportFormat=csv | tee data/morph.csv
```

nanodrop data
```{r, engine='bash', eval=FALSE}

cd ..

curl -L https://docs.google.com/spreadsheets/d/1A8RAq72h2FLstFS7FcUE5hDpYsaJg9ql0M7iGxuN2xA/export?exportFormat=csv | tee data/nanodrop.csv
```

Read in the data to your local R environment.

```{r, eval=TRUE}
setwd('..')
morph <- read.csv(file = "data/morph.csv")  
```

``` {r}
setwd('..')
nano <- read.csv(file = "data/nanodrop.csv")
```

# Cleaning up Data for Morph

```{r, eval=TRUE}
morph$date <- mdy(morph$date) #convert the date column from characters to true date
```

```{r, eval=TRUE}
morph <- morph %>%
  separate(jar_id, c("treatment", "replicate"), sep = "-") #create two new columns, treatment and replicate from jar id 
```

```{r, eval=TRUE}
morph <- morph %>%
  mutate(treatment = as.factor(treatment)) %>%
  mutate(stage = as.factor(stage)) %>%
  mutate(animal_id = as.factor(animal_id)) %>%
  mutate(date = as.factor(date)) %>%
  mutate(treatment_order = factor(paste(treatment, animal_id))) # Create a new variable for ordering by treatment
```

```{r}
# Create a new column 'simple_stage' based on conditions
morph <- morph %>%
  mutate(simple_stage = case_when(
    stage %in% c("A1", "A2") ~ "A",
    stage %in% c("B1", "B2") ~ "B",
    stage %in% c("C1", "C2") ~ "C",
    stage == "TO" ~ "D",
    TRUE ~ NA_character_  # This will handle any other cases or return NA if none match
  ))
```

New data frame with only the rows after 24 hours of exposure when the RNA samples were saved.

```{r}
morph_24 <- morph[morph$hpe == 24, ]
```



# Cleaning up Nano Data

Extract identifiers from the sample_id column and make new columns.

```{r}
nano <- nano %>%
  mutate(jar_id = str_extract(sample_id, "(0|5|45)-R[1-4]"))%>%
  mutate(experiment = str_extract(sample_id, "E[1-4]"))
```

```{r, eval=TRUE}
nano <- nano %>%
  separate(jar_id, c("treatment", "replicate"), sep = "-") #create two new columns, treatment and replicate from jar id 
```


```{r, eval=TRUE}
nano$date <- mdy(nano$date) #convert the date column from characters to true date
```

```{r, eval=TRUE}
nano <- nano %>%
  mutate(treatment = as.factor(treatment)) %>%
  mutate(experiment = as.factor(experiment)) %>%
  mutate(date = as.factor(date))
```

# Join Data Frames

```{r}
# Perform the inner join with an additional condition
merged_data <- inner_join(nano, morph, by = c("date", "experiment", "replicate", "treatment"))
```

# Reduce data frame to only columns of interest

```{r}
merged <- merged_data[, c("date", "rna_conc_ng/uL", "A260.A280", "treatment", "experiment", "replicate", "stage", "stage_2", "simple_stage", "animal_id", "health", "attacment")]
```

